---
title: "HD Age/Sex adjustemnt of 5k"
output:
  html_document:
    df_print: paged
  pdf_document: default
  html_notebook: default
---

THIS SCRIPT performs adjustment for HD age/sex for high signalling somamers in 5kdataset

- All MS and HD samples were used to calculate SNR
- high signaling somamers are then adjusted for HD age and sex
- adjusted somamers for untreated MS samples are then correlated with disability outcomes
- correlation coefficients are submitted to IPA for Core analysis 

###################

Set the environment 

```{r}

remove(list = ls())


library(readr, quietly = TRUE)
library(readxl, quietly = TRUE)
library(dplyr, quietly = TRUE)
library(data.table, quietly = TRUE)
library(foreach, quietly = TRUE)
library(tidyr)
library(ggplot2)
library(stringr)
library(lme4)
library(ggpubr)
library(openxlsx)
library(glmnet)
library(tidyr)
library(broom)
library(REdaS)
library(vioplot)
library(epiR)

```




load functions

```{r}
'%notin%' <- function(x,y)!('%in%'(x,y))


tsapply <- function(...){t(sapply(...))}


remove_outlier <-
  function(x, mult, ...) {
      upper <-
      quantile(x, 3 / 4, na.rm = TRUE) + mult * IQR(x, na.rm = TRUE)   
    downer <-
      quantile(x, 1 / 4, na.rm = TRUE) - mult * IQR(x, na.rm = TRUE)   
       out <- ifelse(x > upper | x < downer, NA, x)
    return(out)
  }

remout_fun <- function(dat, outcomes, mult, ...) {
    dat_noout <- dat %>%
       mutate_at(outcomes, remove_outlier, mult = mult)
  return(dat_noout)
}


curb_outlier <-
  function(x, mult = 3, ...) {
    upper <-
      quantile(x, 3 / 4, na.rm = TRUE) + mult * IQR(x, na.rm = TRUE)   
    downer <-
      quantile(x, 1 / 4, na.rm = TRUE) - mult * IQR(x, na.rm = TRUE)   
   
     out <- ifelse(x > upper, upper, ifelse(x < downer, downer, x))
    
   return(out)
  }



remove_singleton <- function(data){
  if(is.null(groups(data))==TRUE){
    stop("Need to pass a dataset with a grouping structure")
  }
  if(is.null(groups(data))==FALSE){
    newdata <- data %>%
      nest %>%
      mutate(nrows = sapply(data,function(dat){dim(dat)[1]})) %>%
      filter(nrows > 1 ) %>%
      dplyr::select(-nrows) %>% 
      unnest
    return(newdata)
  }
}



```

Load raw data

```{r}
demo_5k <- read_csv("./input/outcomes_5k.csv")
soma_5k <- fread("./input/soma_raw_5k.csv")
transl_5k <- read_excel("./input/translation file 5k new.xlsx")

markers_5k <- transl_5k$SomaID
seqid_5k <- transl_5k$seqid

```
convert somaID to seqID
```{r}
soma_5k <- soma_5k %>% 
  dplyr::select(sampleid,all_of(markers_5k))

names(soma_5k)[-1] <- seqid_5k
```

merge demo and soma
```{r}
data_5k <- merge(demo_5k,soma_5k,by="sampleid")
```

#########################################################
             log-transform data
#########################################################

```{r}
data_5k_log <- data_5k %>% 
  dplyr::mutate_at(seqid_5k,.funs = log)
```

#########################################################
 generate HD cohort and identify outlier samples
 (samples that have > 15% of somamers as outliers)
#########################################################

5K

```{r}
data_5k_hd <- data_5k_log %>% 
  filter(diagnosis=="HD")

```

remove outliers in the 5k HD cohort

```{r}
data_5k_hd_noout <- remout_fun(dat = data_5k_hd,outcomes = seqid_5k,mult = 1.5)

a <- as.data.frame(apply(data_5k_hd_noout[,seqid_5k], MARGIN = 1,function(x) sum(is.na(x))))

```

remove samples with more than 15% of somamers as outliers

```{r}
data_5k_hd_noout <- data_5k_hd_noout[which(a$`apply(data_5k_hd_noout[, seqid_5k], MARGIN = 1, function(x) sum(is.na(x)))`<0.15*length(seqid_5k)),]

```

identify HD outliers in the HD cohort

```{r}
to_remove_5k <- data_5k_hd$sampleid[which(data_5k_hd$sampleid %notin% data_5k_hd_noout$sampleid)]

hist(a$`apply(data_5k_hd_noout[, seqid_5k], MARGIN = 1, function(x) sum(is.na(x)))`,main = "# of soma as outliers per sample",xlab = "# outliers")

```
remove HD outliers from whole data

```{r}
data_5k_log <- data_5k_log[-which(data_5k_log$sampleid %in% to_remove_5k),]
```


###############################################################################

 calculate SNR in selected somamers - using longitudinal HD samples 
 w/o outliers as technical replicates 
 and MS cohort samples to calculate biological variance 

###############################################################################

5k

```{r}
techrep_potential <- data_5k_hd_noout %>% 
  group_by(patientcode) %>% 
  remove_singleton() %>% 
  ungroup() %>% 
  arrange(patientcode) 

training <- data_5k_log %>% 
  filter(diagnosis %in% c("RR-MS","SP-MS","PP-MS"))
```


CALCULATE SNR

```{r}
marker_tibble <- tibble::enframe(seqid_5k,name = NULL,value = "seqid")

marker_tibble$tech_sigma <- NA
marker_tibble$clin_sigma <- NA


for(i in 1:length(seqid_5k)){
  # i <- 1
  

    
  f <- as.formula(paste0("`",seqid_5k[i],"`","~ (1|patientcode)"))
  
  
  tech_var <- lmer(formula = f,data = techrep_potential )
  
  marker_tibble$tech_sigma[i] <- summary(tech_var)$sigma
  
  fr <- as.formula(paste0("`",seqid_5k[i],"`"," ~ 1"))
  
  
  clin_var <- lm(formula = fr,data = training )
  
  
  marker_tibble$clin_sigma[i] <- summary(clin_var)$sigma
  
}


# calculate SNR
var_comps <- marker_tibble %>% 
  mutate(snr=clin_sigma^2/(clin_sigma^2+tech_sigma^2)) 

# sort SNR
var_comps <- var_comps %>% 
  arrange(desc(snr))

# hist(var_comps$snr)

var_comps_5k_all <- var_comps


r <- var_comps_5k_all %>% 
  ggplot(aes(x=snr)) +  
  geom_histogram(fill="white",color="blue",bins = 50) +
  xlab("SNR")+
  # xlim(c(0,1))+
  # ylim(c(0,260))+
  geom_vline(xintercept = 0.8, color="red")+
  ggtitle(paste0("5k SOMA SNR (80% cut-off: ", length(which(var_comps_5k_all$snr>0.8))," soma retained)")) +
  theme_bw()+
  theme(
    # legend.background = element_rect(fill="transparent"),
    axis.text.x = element_text(size=12,colour="#404040",hjust = 1, vjust = 1,angle = 45),
    axis.text.y = element_text(size=12,hjust = 0.5,vjust = .5 ),
    axis.title.y = element_text(color="#404040",size=14),
    axis.title.x = element_text(color="#404040",size=14),
    # legend.box.background = element_rect(colour = "black",fill="white"),
    legend.position = "right",
    axis.line = element_blank(),
    axis.line.x = element_blank(),
    axis.line.y = element_blank(),
    # panel.grid.major.y = element_line(color="gray",size=.25),
    panel.border = element_rect(colour = "black", fill=NA, size=0.5)
  )
r


write_csv(var_comps_5k_all,"./output/age_sex_adj/5k_SNR.csv")
```

Isolate somamers that passed 0.8 SNR cutoff


```{r}
var_selection <- var_comps_5k_all %>% 
  filter(snr>=0.8)

#merge with translation file 

var_selection <- left_join(var_selection,transl_5k)

print(paste0("Number of high-SNR somamers: ", length(var_selection$seqid)))

high_markers <- var_selection$seqid
high_markers_soma <- var_selection$SomaID
```

#############################################################################################

 isolation of somamers significantly correlated with age and gender in the INTERVAL study

#############################################################################################

```{r}
pub_dat <- read_excel("./input/Genomic atlas of serum proteins Tables.xlsx",sheet=2,skip=4,
                      col_names = FALSE,na=c("",NA,"--"))
pub_dat <- pub_dat[,-c(19)]

names(pub_dat) <- c("id","protein","target","uniprot","r2_adj","sub_pass","age_beta_pub","age_se_pub","age_pval_pub",
                    "female_beta_pub","female_se_pub","female_pval_pub","bmi_beta_pub","bmi_se_pub","bmi_pval_pub",
                    "egfr_beta_pub","egfr_se_pub","egfr_pval_pub")

pub_dat <- pub_dat %>% 
  mutate(age_pval_pub = 10^(-1*age_pval_pub),
         female_pval_pub = 10^(-1*female_pval_pub)) %>% 
  select(id:female_pval_pub)

```


load a table of 5k Somamers

```{r}
translate_file <- read_excel("./input/translation file 5k new.xlsx")

# filter high SNR Somamers, and generate "id" column
tomatch <- translate_file %>% 
  filter(seqid %in% high_markers) %>% 
  dplyr::rename(gene = Gene,
         protein = Target,
         protein_full=Target_description,
         marker = SomaID)%>% 
  mutate(id = marker) %>% 
  mutate(id = str_replace_all(id,"-",".")) %>% 
  mutate(id = str_replace_all(id,"_",".")) %>% 
  mutate(id = str_replace_all(id,",",".")) %>%
  mutate(id = str_replace_all(id,"@",".")) %>%
  mutate(id = str_replace_all(id," ",".")) %>%
  mutate(id = str_replace_all(id,"(\\.)(\\.)",".")) 

# merge INTERVAL markers with 5k markers
matched_markers <- tomatch$marker[match(pub_dat$id, tomatch$id)]
matched_markers <- unique(matched_markers[!is.na(matched_markers)])


#convert to seqid
matched_markers <- tibble::enframe(matched_markers,name = NULL,value = "SomaID")
matched_markers <- left_join(matched_markers,translate_file[,1:2])
matched_markers <- matched_markers$seqid


tomatch <- tomatch %>% 
  filter(seqid %in% matched_markers)


# isolate INTERVAL proteins that correlate with age with p-value <1e-5
age_proteins <- pub_dat$id[which(pub_dat[["age_pval_pub"]]<1e-5)]
# filter only proteins present in 5k
prev_age <- with(tomatch,marker[id %in% age_proteins])

#convert to seqid
prev_age <- tibble::enframe(prev_age,name = NULL,value = "SomaID")
prev_age <- left_join(prev_age,translate_file[,1:2])
prev_age <- prev_age$seqid

# isolate INTERVAL proteins that correlate with sex with p-value <1e-5
sex_proteins <- pub_dat$id[which(pub_dat[["female_pval_pub"]]<1e-5)]
# filter only proteins present in 5k
prev_sex <- with(tomatch,marker[id %in% sex_proteins])

#convert to seqid
prev_sex <- tibble::enframe(prev_sex,name = NULL,value = "SomaID")
prev_sex <- left_join(prev_sex,translate_file[,1:2])
prev_sex <- prev_sex$seqid

```



###############################################################
Only Healthy Donors
###############################################################

```{r}
# get coefficients for age and gender in HD cohort
hd_results <- tsapply(matched_markers,function(char){
    mod <- lm(data_5k_hd_noout[[char]]~age+gender,data=data_5k_hd_noout)
  out <- c(as.numeric(summary(mod)$coefficients[2,c(1,4)]),as.numeric(summary(mod)$coefficients[3,c(1,4)]))
  
  if(char %notin% prev_age){out[c(1:2)] <- NA}
  if(char %notin% prev_sex){out[c(3:4)] <- NA}
  
  names(out) <- c("age_beta_hd","age_pval_hd","female_beta_hd","female_pval_hd")
  return(out)
})

hd_results <- as_tibble(hd_results) %>%
  mutate(seqid = matched_markers) %>%
  select(seqid,everything()) %>%
  mutate(age_pval_hd = p.adjust(age_pval_hd,method="none")) %>%
  mutate(female_pval_hd = p.adjust(female_pval_hd,method="none")) 

hd_results <- inner_join(tomatch,hd_results,by="seqid") %>% 
  select(id,marker,gene,protein,protein_full,everything()) 
```



#########################################
  Compare to INTERVAL study
#########################################

```{r}
pub_dat <- pub_dat %>% 
  select(-protein,-target,-uniprot)

new_results <- inner_join(hd_results,pub_dat,by="id")

```


isolate markers that are significant (p<0.05, no multiple comparison adjustment) in NDS HD cohort, significant in INTREVAL (p<1e-5) and have the same directionality

```{r}
age_to_adjust <- with(new_results,marker[which(age_pval_hd <= 0.05 & seqid %in% prev_age & sign(age_beta_hd)*sign(age_beta_pub)==1)])
sex_to_adjust <- with(new_results,marker[which(female_pval_hd <= 0.05 & seqid %in% prev_sex & sign(female_beta_hd)*sign(female_beta_pub)==1)])
```


#############################################################################################

 isolation of somamers significantly correlated with age and gender in the CHIANTI study

#############################################################################################

```{r}
pub_dat2 <- read_csv("./input/chianti_data.csv")


translate_file <- read_excel("./input/translation file 5k new.xlsx")

tomatch2 <- translate_file %>% 
  filter(seqid %in% high_markers) %>% 
  dplyr::rename(gene = Gene,
         protein = Target,
         protein_full=Target_description,
         marker = SomaID) %>% 
  mutate(id = marker) %>% 
  mutate(id = str_replace_all(id,"-",".")) %>% 
  mutate(id = str_replace_all(id,"_",".")) %>% 
  mutate(id = str_replace_all(id,",",".")) %>%
  mutate(id = str_replace_all(id,"@",".")) %>%
  mutate(id = str_replace_all(id," ",".")) %>%
  mutate(id = str_replace_all(id,"(\\.)(\\.)",".")) 

```


merge CHIANTI markers with 5k markers

```{r}
matched_markers2 <- tomatch2$marker[match(pub_dat2$id, tomatch2$id)]
matched_markers2 <- unique(matched_markers2[!is.na(matched_markers2)])

tomatch2 <- tomatch2 %>% 
  filter(marker %in% matched_markers2)

#convert to seqid
matched_markers2 <- tibble::enframe(matched_markers2,name = NULL,value = "SomaID")
matched_markers2 <- left_join(matched_markers2,translate_file[,1:2])
matched_markers2 <- matched_markers2$seqid
```


isolate CHIANTI proteins that correlate with age with p-value <5e-2
```{r}
age_proteins <- pub_dat2$id[which(pub_dat2[["age_pval_pub2"]]<5e-2)]

# filter only proteins present in 5k

prev_age2 <- with(tomatch2,marker[id %in% age_proteins])


#convert to seqid
prev_age2 <- tibble::enframe(prev_age2,name = NULL,value = "SomaID")
prev_age2 <- left_join(prev_age2,translate_file[,1:2])
prev_age2 <- prev_age2$seqid

# isolate CHIANTI proteins that correlate with sex with p-value <5e-2
sex_proteins <- pub_dat2$id[which(pub_dat2[["sex_pval_pub2"]]<5e-2)]

# filter only proteins present in 5k
prev_sex2 <- with(tomatch2,marker[id %in% sex_proteins])

#convert to seqid
prev_sex2 <- tibble::enframe(prev_sex2,name = NULL,value = "SomaID")
prev_sex2 <- left_join(prev_sex2,translate_file[,1:2])
prev_sex2 <- prev_sex2$seqid
```


###############################################################
 Only Healthy Donors 
###############################################################

get coefficients for age and gender in HD cohort

```{r}
hd_results2 <- tsapply(matched_markers2,function(char){
  mod <- lm(data_5k_hd_noout[[char]]~age+gender,data=data_5k_hd_noout)
  out <- c(as.numeric(summary(mod)$coefficients[2,c(1,4)]),as.numeric(summary(mod)$coefficients[3,c(1,4)]))
  
  if(char %notin% prev_age2){out[c(1:2)] <- NA}
  if(char %notin% prev_sex2){out[c(3:4)] <- NA}
  
  names(out) <- c("age_beta_hd","age_pval_hd","female_beta_hd","female_pval_hd")
  return(out)
})

hd_results2 <- as_tibble(hd_results2) %>%
  mutate(seqid = matched_markers2) %>%
  select(seqid,everything()) %>%
  mutate(age_pval_hd = p.adjust(age_pval_hd,method="none")) %>%
  mutate(female_pval_hd = p.adjust(female_pval_hd,method="none")) 

hd_results2 <- inner_join(tomatch2,hd_results2,by="seqid") %>% 
  select(id,marker,gene,protein,protein_full,everything()) 
```



#########################################
 Compare to CHIANTI 
#########################################

```{r}
new_results2 <- inner_join(hd_results2,pub_dat2,by="id")
```


isolate markers that are significant (p<0.05, no multiple comparison adjustment) in NDS HD cohort, significant in CHIANTI (p<0.05) and have the same directionality

```{r}
age_to_adjust2 <- with(new_results2,marker[which(age_pval_hd <= 0.05 & seqid %in% prev_age & sign(age_beta_hd)*sign(age_beta_pub2)==1)])
sex_to_adjust2 <- with(new_results2,marker[which(female_pval_hd <= 0.05 & seqid %in% prev_sex & sign(female_beta_hd)*sign(sex_beta_pub2)==1)])

```


merge with previous and isolate unique
```{r}
age_to_adjust <- unique(c(age_to_adjust,age_to_adjust2))
sex_to_adjust <- unique(c(sex_to_adjust,sex_to_adjust2))

```


########################################################################
  Additional HD results 
########################################################################

look at markers not already selected by INTREVAL and CHIANTI cohort 

```{r}
age_other_markers <- high_markers_soma[high_markers_soma %notin% age_to_adjust]
sex_other_markers <- high_markers_soma[high_markers_soma %notin% sex_to_adjust]
other_markers <- unique(c(age_other_markers,sex_other_markers))

#convert to seqid
other_markers <- tibble::enframe(other_markers,name = NULL,value = "SomaID")
other_markers <- left_join(other_markers,translate_file[,1:2])
other_markers <- other_markers$seqid

age_other_markers <- tibble::enframe(age_other_markers,name = NULL,value = "SomaID")
age_other_markers <- left_join(age_other_markers,translate_file[,1:2])
age_other_markers <- age_other_markers$seqid

sex_other_markers <- tibble::enframe(sex_other_markers,name = NULL,value = "SomaID")
sex_other_markers <- left_join(sex_other_markers,translate_file[,1:2])
sex_other_markers <- sex_other_markers$seqid
```


get coefficients 

```{r}
hd_other_results <- tsapply(other_markers,function(char){
  mod <- lm(data_5k_hd_noout[[char]]~age+gender,data=data_5k_hd_noout)
  out <- c(as.numeric(summary(mod)$coefficients[2,c(1,4)]),as.numeric(summary(mod)$coefficients[3,c(1,4)]))
  
  if(char %notin% age_other_markers){out[c(1:2)] <- NA}
  if(char %notin% sex_other_markers){out[c(3:4)] <- NA}
  
  names(out) <- c("age_beta_hd","age_pval_hd","female_beta_hd","female_pval_hd")
  return(out)
})

# adjust for multiple comparisons by running FDR adjustment 
hd_other_results <- as_tibble(hd_other_results) %>%
  mutate(marker = other_markers) %>%
  select(marker,everything()) %>%
  mutate(age_pval_hd = p.adjust(age_pval_hd,method = "fdr" )) %>%
  mutate(female_pval_hd = p.adjust(female_pval_hd,method="fdr"))


trans_file <- translate_file
names(trans_file)[2] <- "marker"
names(trans_file) <- tolower(names(trans_file))

hd_other_results <- inner_join(trans_file,hd_other_results,by="marker") %>%
  select(somaid,marker,gene,target,target_description,everything())

new_age <- with(hd_other_results,somaid[age_pval_hd<=0.05 & !is.na(age_pval_hd)])
new_sex <- with(hd_other_results,somaid[female_pval_hd<=0.05 & !is.na(female_pval_hd)])

print(paste0("new age: ",length(new_age), " & new sex: ",length(new_sex)))

```

merge with previously identified markers 

```{r}
age_to_adjust <- c(age_to_adjust,new_age)
sex_to_adjust <- c(sex_to_adjust,new_sex)

# convert to seqid
age_to_adjust <- tibble::enframe(age_to_adjust,name = NULL,value = "SomaID")
age_to_adjust <- left_join(age_to_adjust,translate_file[,1:2])
age_to_adjust <- age_to_adjust$seqid

sex_to_adjust <- tibble::enframe(sex_to_adjust,name = NULL,value = "SomaID")
sex_to_adjust <- left_join(sex_to_adjust,translate_file[,1:2])
sex_to_adjust <- sex_to_adjust$seqid


print(paste0("# age to adjust: ",length(age_to_adjust), " & # sex to adjust: ",length(sex_to_adjust)))
```

```{r}
writeLines(age_to_adjust,"./input/age_to_adjust.txt")
writeLines(sex_to_adjust,"./input/sex_to_adjust.txt")
```



run lm in HD cohort for all somamers

```{r}
hd_results_to_save <- tsapply(high_markers,function(char){
  mod <- lm(data_5k_hd_noout[[char]]~age+gender,data=data_5k_hd_noout)
  out <- c(as.numeric(summary(mod)$coefficients[2,c(1,4)]),as.numeric(summary(mod)$coefficients[3,c(1,4)]))
  
  names(out) <- c("age_beta_hd","age_pval_hd","female_beta_hd","female_pval_hd")
  return(out)
})

# adjust for multiple comparisons by running FDR adjustment 
hd_results_to_save <- as_tibble(hd_results_to_save) %>%
  mutate(marker = high_markers) %>%
  select(marker,everything()) %>%
  mutate(age_pval_hd = p.adjust(age_pval_hd,method = "fdr" )) %>%
  mutate(female_pval_hd = p.adjust(female_pval_hd,method="fdr"))
 

```



######################################################
adjust for age and sex using significant somamers
######################################################

```{r}
adjust_function <- function(y_char,data=new,alpha_val=0.05,...){
  # y_char <- "SL000001/SL000002"
  # y_char <- "SL000001"
  # data <- new
  y<-data[[y_char]]
  if(any(str_detect(y_char,c(age_to_adjust,sex_to_adjust)))){
    
    if(any(str_detect(y_char,age_to_adjust)) & any(str_detect(y_char,sex_to_adjust))){
      mod <- lm(y~age+gender,data)
    }else{
      if(any(str_detect(y_char,age_to_adjust))){
        mod <- lm(y~age,data)
      }
      if(any(str_detect(y_char,sex_to_adjust))){
        mod <- lm(y~gender,data)
      }
    }
  }else{
    mod <- lm(y~1)
  }
  return(mod)
}

soma_new <- data_5k_log

soma_new_adj <- foreach(i=1:length(high_markers),.final=as.data.table,.packages=c("dplyr")) %do% {
  mod <- adjust_function(high_markers[i],data=data_5k_hd_noout)
  (soma_new[[high_markers[i]]] - predict(mod,newdata=soma_new))
}

names(soma_new_adj) <- high_markers
soma_new_adj[,sampleid := soma_new$sampleid]
setcolorder(soma_new_adj,c("sampleid"))


```


#############################################
#############################################
    scale to mean 0 and sd of 1 in HD
#############################################
#############################################

```{r}
# isolate HD data only

soma_adj_demo <- merge(demo_5k,soma_new_adj,all.y=TRUE)

data_hd <- soma_adj_demo %>% 
  filter(diagnosis=="HD")

# remove outliers in the HD cohort

data_hd_noout <- remout_fun(dat = data_hd,outcomes = high_markers,mult = 1.5)

a <- as.data.frame(apply(data_hd_noout[,high_markers], MARGIN = 1,function(x) sum(is.na(x))))

# get means and sd of HD
means <- apply(data_hd_noout[,high_markers],MARGIN = 2,FUN = function(x){mean(x,na.rm=TRUE)})
stdev <- apply(data_hd_noout[,high_markers],MARGIN = 2,FUN = function(x){sd(x,na.rm=TRUE)})

data_scaled <- soma_adj_demo %>% 
  dplyr::select(-all_of(high_markers))

for(i in 1:length(high_markers)){
  # i <- 1
  data_scaled[,high_markers[i]] <- (soma_adj_demo[,high_markers[i]] - means[i]) / stdev[i]
}

#check if mean is 0 and sd is 1:

data_hd_new <- data_scaled %>% 
  filter(diagnosis=="HD")

means <- apply(data_hd_new[,high_markers],MARGIN = 2,FUN = mean)
stdev <- apply(data_hd_new[,high_markers],MARGIN = 2,FUN = sd)


```


adjust outliers above and below 3*IQR to the value of the outlier cutoff value


```{r}
data_noout_curb <- curb_fun(dat = data_all,outcomes = vars,mult = 3)

# save files 

write_csv(data_noout_curb,"./input/soma_highSNR_adj_scaled_3_iqr_curbed.csv")

```

