---
title: "Calculation of pathway scores that differemtiates MS from HD in 5k dataset"
output:
  html_document:
    df_print: paged
  pdf_document: default
  html_notebook: default
---

THIS SCRIPT calculates pathway scores for MS vs HD difference

_ only pathways with -log10(pval)>5 were considered

- MS vs HD difference sign used for directionality 

- scores calculated using HD age/sex-adjusted and HD scaled somamers (3714 somamers - human only) with outliers CURBED


###################

Set the environment 

```{r}

remove(list = ls())


library(readr, quietly = TRUE)
library(readxl, quietly = TRUE)
library(dplyr, quietly = TRUE)
library(data.table, quietly = TRUE)
library(foreach, quietly = TRUE)
library(tidyr)
library(ggplot2)
library(stringr)
library(lme4)
library(ggpubr)
library(openxlsx)
library(glmnet)
library(tidyr)
library(broom)
library(REdaS)
library(vioplot)
library(epiR)
library(broom)
library(MatchIt)
library(gridExtra)
library(rstatix)

```


load functions

```{r}
'%notin%' <- function(x,y)!('%in%'(x,y))


```

Load raw data

```{r}

# load vector of high-signaling somamers

high_markers <- readLines("./input/high_SNR_markers.txt")


# load curbed dataset

data <- fread("./input/soma_highSNR_adj_scaled_3_iqr_curbed.csv")




# select only somamers of interest and only untreated MS and HD samples

data <- data %>% 
  # filter(nice_therapy=="Untreated") %>% 
   filter(diagnosis %in% c("HD","RR-MS","SP-MS","PP-MS"))

table(data$diagnosis)




demo_5k <- read_csv("./input/outcomes_5k.csv")
transl_5k <- read_excel("./input/translation file 5k new.xlsx")


markers_5k <- transl_5k$SomaID
seqid_5k <- transl_5k$seqid

```


```{r}

# isolate samples of interest

samples <- data$sampleid

# isolate somamers data only 

soma_scaled <- data %>% 
  dplyr::select(sampleid,all_of(high_markers))

# transpose the data

soma_scaled_t <- as.data.frame(t(as.matrix(soma_scaled)))

names(soma_scaled_t) <- soma_scaled_t[1,]
soma_scaled_t <- soma_scaled_t[-1,]

soma_scaled_t$somaid <- rownames(soma_scaled_t)

rownames(soma_scaled_t) <- NULL

soma_scaled_t <- soma_scaled_t %>% 
  select(somaid,everything()) 

soma_scaled_t <- soma_scaled_t %>% 
  mutate_at(.vars = names(soma_scaled_t)[-1],.funs = as.numeric)

class(soma_scaled_t$`S-181218-03143`)

soma_selection <- soma_scaled_t %>% 
  select(somaid,all_of(samples))



```


################################################################################
################################################################################
#
#  get somamers for each pathways
#
################################################################################
################################################################################

```{r}
# load pathway data

pathway_data <- read_excel("./input/IPA_pathways_all.xlsx")

# isolate pathways for GMS disability outcome and filter pathwyas with -log10(pval) > 5 

pathway_data <- pathway_data %>% 
  filter(outcome=="Disease effects: MS vs HD") %>% 
  filter(pval>5)

# generate vector of unique pathways

pathways <- sort(unique(pathway_data$pathway))


# load correlation of somamers with GMSdis outcome

  corr <- read_csv("./Figures/IPA_analysis_files/ms_vs_hd_forIPA_selection.csv")
  
  corr$assigned_directionality <- ifelse(corr$ms_hd_median_diff<0,yes = -1,no = 1)

  names(corr)[1] <- "somaid"
  
  # merge with soma data

soma_selection <- left_join(corr,soma_selection,by="somaid")
```


############################
#########################
MS vs HD PATHWAY SCORES calculation
#########################
############################

```{r}
# get a dataframe of pathways for progression score

results <- tibble::enframe(x = samples,name = NULL,value = "sampleid")

pathway_soma <- tibble::enframe(soma_selection$somaid,name = NULL,value = "somaid")

which(pathways=="Wound Healing Signaling Pathway")

nr_soma <- vector()

cor_dir_sum <- vector()

# generate a file for supplementary table - summary of pathways  

summary_pathways <- data.frame(matrix(nrow = 0,ncol = 6))

names(summary_pathways) <- c("pathway","type","method","seqid","gene","coefficient")

for(j in 1:length(pathways)){
# j <- 1
  
selection <- pathway_data %>% 
    filter(pathway==pathways[j])
  
# get a vector of markers
markers <- strsplit(selection$molecules,"[,/ -()]")[[1]]

  # isolate required data
  
  final_data <- soma_selection %>% 
    filter(gene %in% markers)
  
 
  # final_data[,samples] <- final_data$assigned_directionality * final_data[,samples]
  
  final_data[,samples] <- final_data$ms_hd_median_diff * final_data[,samples]
  
  cor_dir_sum <- c(cor_dir_sum,sum(final_data$ms_hd_median_diff))
  
  nr_somamers <- length(final_data$somaid)
  
  # calculate the score
  
  score <- apply(final_data[,samples],MARGIN = 2,FUN = sum, na.rm=TRUE)
  # score <- apply(final_data[,samples],MARGIN = 2,FUN = mean, na.rm=TRUE)
  
  # mutliply score by sign of the sum of median differences between MS and HD for the pathway
  
  cor_dir <- ifelse(sum(final_data$ms_hd_median_diff)>0,yes = 1,no = -1)
  
  score <- score * cor_dir
  # score <- score / nr_somamers
  results[[paste0("pathway_",j,"_score")]] <- score  
  
  # selected somamers
  pathway_soma[[paste0("pathway_",j,"_score")]] <- NA
  
  pathway_soma[which(pathway_soma$somaid %in% final_data$somaid),j+1] <- 1
  pathway_soma[which(pathway_soma$somaid %notin% final_data$somaid),j+1] <- 0

  nr_soma <- c(nr_soma,nr_somamers)
  
  #for summary
  for_summary <- final_data %>% 
    mutate(pathway=pathways[j]) %>% 
    mutate(type="ms vs hd") %>% 
    mutate(method="IPA") %>% 
    dplyr::select(pathway,type,method,somaid,gene,ms_hd_median_diff) %>% 
    dplyr::rename(coefficient=ms_hd_median_diff)
  
  summary_pathways <- rbind(summary_pathways,for_summary)

      
}

nr_soma_for_scores <- tibble::enframe(pathways,name = NULL,value = "pathway")

nr_soma_for_scores$nr_soma <- nr_soma

write_csv(nr_soma_for_scores,"./output/MS_vs_HD/nr_soma_in_scores_ms_vs_hd.csv")

coef_data <- tibble::enframe(pathways,name = NULL,value = "pathway")

# coef_data$GMSdis_Rho_mean <- cor_dir_mean
coef_data$MS_vs_HD_median_diff <- cor_dir_sum


final_scores_tosave <- results

names(final_scores_tosave)[which(grepl("pathway",names(final_scores_tosave)))] <- pathways


write_csv(final_scores_tosave,"./output/MS_vs_HD/ms_vs_hd_pathways_scores.csv")

# save soma selection in pathways 
names(pathway_soma)[which(grepl("pathway",names(pathway_soma)))] <- pathways

names(pathway_soma)[1] <- "seqid"

pathway_soma <- right_join(transl_5k[,-1],pathway_soma)

write_csv(pathway_soma,"./output/MS_vs_HD/soma_selection_in_ms_vs_hd_pathways.csv")

# save summary pathways 

write_csv(summary_pathways,"./output/MS_vs_HD/ms_vs_hd_summary_pathways.csv")
```
