---
title: "Generation of new non-IPA pathway in 5k dataset for somamers that differentiate MS from HD"
output:
  html_document:
    df_print: paged
  pdf_document: default
  html_notebook: default
---

THIS SCRIPT identifies somamers not used in any of the IPA pathways for MS vs HD and uses 
correlations to group them and generate new unannotated pthways 

- using HD age/sex-adjusted and HD scaled and CURBED somamers (3714 somamers)


###################

Set the environment 

```{r}

remove(list = ls())


library(readr, quietly = TRUE)
library(readxl, quietly = TRUE)
library(dplyr, quietly = TRUE)
library(data.table, quietly = TRUE)
library(foreach, quietly = TRUE)
library(tidyr)
library(ggplot2)
library(stringr)
library(lme4)
library(ggpubr)
library(openxlsx)
library(glmnet)
library(tidyr)
library(broom)
library(REdaS)
library(vioplot)
library(epiR)
library(broom)
library(MatchIt)
library(gridExtra)
library(rstatix)
library(ggcorrplot)

```


load functions

```{r}
'%notin%' <- function(x,y)!('%in%'(x,y))


tsapply <- function(...){t(sapply(...))}



 
cor_est <- function(x,y,...){
  obj <- cor.test(x,y,...)
  val <- obj$estimate
  return(round(as.numeric(val),digits = 3))
}
cor_pval <- function(x,y,...){
  obj <- cor.test(x,y,...)
  pval <- obj$p.value
  return(as.numeric(pval))
}


corr_fun <- function(dat,outcomes,high_markers,...){
    dat_est <- sapply(outcomes,function(ychar){
    sapply(high_markers,function(xchar){
      cor_est(dat[[xchar]],dat[[ychar]],method="spearman")
    })
  })
  
 
  dat_pval <- sapply(outcomes,function(ychar){sapply(high_markers,function(xchar){
    cor_pval(dat[[xchar]],dat[[ychar]],method="spearman")})})
  
  
  dat_pval <- as_tibble(dat_pval) %>% 
    mutate_all(p.adjust,method="fdr") %>%
    mutate(marker = high_markers,type="p-value") %>%  
    dplyr::select(marker,type,everything()) 
  
  dat_est <- as_tibble(dat_est) %>% 
    mutate(marker = high_markers,type="spearman") %>% 
    dplyr::select(marker,type,everything())
  
  dat_results <- bind_rows(dat_est,dat_pval) # stack em up
  dat_results <- dat_results %>% 
    gather(key,value,outcomes) 
 
  
  dat_results <- dat_results %>% 
    mutate(key = factor(key,levels=unique(dat_results$key))) %>% 
    mutate(type=factor(type,levels=c("spearman","p-value"))) %>% 
    arrange(marker,key,type) %>%    
    mutate(key = paste(as.character(key),type)) %>% 
    dplyr::select(-type) 
  
  dat_results <- dat_results %>% 
    mutate(key = factor(key,levels=unique(dat_results$key))) %>% 
    spread(key,value) 
  
  
  
  return(dat_results)
}




```

Load raw data

```{r}
demo_5k <- read_csv("./input/outcomes_5k.csv")
transl_5k <- read_excel("./input/translation file 5k new.xlsx")

soma_scaled <- fread("./input/soma_highSNR_adj_scaled_3_iqr_curbed.csv")

high_markers <- readLines("./input/high_SNR_markers.txt")



markers_5k <- transl_5k$SomaID
seqid_5k <- transl_5k$seqid

```


filter untreated MS samples only
```{r}

# isolate untreated MS samples only 

ms_only <- soma_scaled %>% 
  filter(diagnosis %in% c("RR-MS","PP-MS","SP-MS")) %>%
  filter(nice_therapy=="Untreated")

```

################################################################################
################################################################################
#
#  get somamers for each pathways
#
################################################################################
################################################################################

```{r}
# load pathway data

pathway_data <- read_csv("./input/ipa_ms_vs_hd_pathways_soma_summary.csv")

# isolate used somamers

ipa_markers <- pathway_data %>% 
  filter(nr_pathways>0)
ipa_markers <- ipa_markers$seqid

# load MS_vs_HD differences

  corrs <- read_csv("./input/ms_vs_hd_forIPA_selection.csv")

# isolate somamers not used in IPA scores

not_used <- corrs %>% 
  filter(marker %notin% ipa_markers)

not_used <- not_used$marker


```

GENERATE NEW PATHWAYS from unused somamers

```{r}
# generate a correlation matrix of these
working_data <- ms_only %>% 
  dplyr::select(all_of(not_used))

corr <- cor(working_data,method = "spearman")
corr[(corr)==1] <- NA

# generate histogram of Rho
new_vector <- tibble::enframe(c(corr),name = NULL,value = "cor")
med <- median(new_vector$cor,na.rm=TRUE)

# remove all correlations below abs(0.7)
new_vector_clean <- new_vector %>% 
  filter(!is.na(cor)) %>% 
  filter(abs(cor)>=0.7)
med_clean <- median(new_vector_clean$cor,na.rm=TRUE)



#################################
##
#  generate new pathways
##
#################################

# turn correlation matrix into dataframe
mat <- as_tibble(corr)
# give it "seqid" column
mat$seqid <- names(mat)

# generate a dataframe with all somamers to new pathways
newcluster_results <- tibble::enframe(not_used,name = NULL,value = "seqid")

used_markers <- vector()

for(i in 1:length(not_used)){
  # i <- 1
  # isolate working markers that have not been used previosuly
  markers_iter <- not_used[which(not_used%notin%used_markers)]
  
  # generate correlation matrix with desired set of somamers (all that have 
  # not been used for previous iteration)  
  mat_working <- mat %>% 
    filter(seqid %in% markers_iter) %>% 
    dplyr::select(all_of(markers_iter),seqid)
  
  ##############
  # identify top correlating pair 
  
  # identify maximum value in each column 
  top <- mat_working %>% 
    dplyr::summarise_at(., .vars = markers_iter,.funs = function(x){max(abs(x),na.rm=TRUE)})
  
  # turn it into dataframe
  top <- tibble::enframe((t(as.matrix(top))),name = "marker",value = "max_rho")
  
  # arrange based on descending rho value and isolate top two somamers - 
  # these are the seed for the loop
  top <- top %>%
    dplyr::arrange(desc(max_rho)) %>% 
    slice_head(n=2)
  
  # check if the top correlation still passes the 0.7 cutoff, 
  # if not, break the loop
  if(abs(top$max_rho[1])<=0.7){
        break
  }
  
  # turn them into a vector
  top <- top$marker
  
  # remove them from the working markers 
  working_markers <- markers_iter[-which(markers_iter%in%top)]
  
  # loop through working markers to identify those that correlate 
  # with the rest of the marker with cut-off > abs(0.5)
  
  for(k in 1:length(working_markers)){
    # k <- 1
    # isolate top markers 
    to_test <- mat_working %>% 
      filter(seqid%in%top) %>% 
      dplyr::select(-all_of(top))
    
    # length(names(to_test)[-length(names(to_test))])
    
    # sum up abs rho for all somamers
    sum_rho <- to_test %>% 
      dplyr::summarise_at(.vars = names(to_test)[-length(names(to_test))],.funs = function(x){sum(abs(x),na.rm = TRUE)})
    
    # turn into dataframe
    sum_rho_t <- as_tibble(t(sum_rho))
    # rename column 
    names(sum_rho_t) <- "rho"
    # add column with seid
    sum_rho_t$seqid <- names(sum_rho)
    
    # arrange by descending Rho
    sum_rho_t <- sum_rho_t %>% 
      dplyr::arrange(desc(abs(rho)))
    
    # isolate the top somamer
    new_rho <- sum_rho_t$seqid[1]
    
    # test its correlation with the rest of the somamers - if it passes 
    # cutoff - the number remains, if it's below cut-off, it's replaced by NA
    to_test_na <- to_test %>% 
      dplyr::mutate_at(.vars = names(to_test)[-length(names(to_test))],.funs = function(x) {ifelse(test = abs(x)<=0.7,yes = NA,no = x)})
    
    # test whether new somamer contains NAs
    
    
    if(sum(which(is.na(to_test_na[[new_rho]])))>0){
      top <- top
      break}
    else{
      top <- c(top,new_rho)
      
    }
  }
  
  # add selected markers on the list of used markers
  used_markers <- c(used_markers,top)
  
  # add results to summary dataframe
  to_add <- tibble::enframe(top,name = NULL,value = "seqid")
  to_add[[paste0("New_pathway_",i)]] <- 1
  
  newcluster_results <- left_join(newcluster_results,to_add,by="seqid")
  
  print(i)
  print(length(markers_iter))
  # eliminate pathways with less than 5 somamers
  
  to_eliminate <- tibble::enframe(apply(newcluster_results[,-1],MARGIN = 2,FUN = sum, na.rm=TRUE),name = "pathway",value = "nr_soma")
  
  # filter pathways with less than 5
  to_eliminate <- to_eliminate %>% 
    filter(nr_soma<5)
  to_eliminate <- to_eliminate$pathway
  
  #remove these pathways from the result dataframe
  newcluster_results <- newcluster_results %>% 
    dplyr::select(-all_of(to_eliminate))
}

# merge with translation file 

newcluster_results <- merge(transl_5k[,1:3],newcluster_results,all.y = TRUE,by="seqid")




```


GENERATE PATHWAY SCORES for unannotated pathways 

```{r}
# isolate samples of interest

samples_interest <- soma_scaled %>% 
  # filter(nice_therapy=="Untreated") %>% 
  filter(diagnosis_simple %in% c("RR-MS","prog-MS","HD"))
  

samples <- samples_interest$sampleid


# transpose soma data

data_5k <- soma_scaled %>% 
  dplyr::select(sampleid,all_of(high_markers))
 
soma_scaled_t <- as.data.frame(t(as.matrix(data_5k)))

names(soma_scaled_t) <- soma_scaled_t[1,]
soma_scaled_t <- soma_scaled_t[-1,]

soma_scaled_t$somaid <- rownames(soma_scaled_t)

rownames(soma_scaled_t) <- NULL


soma_scaled_t <- soma_scaled_t %>% 
  select(somaid,everything()) 

soma_scaled_t <- soma_scaled_t %>% 
  mutate_at(.vars = names(soma_scaled_t)[-1],.funs = as.numeric)

class(soma_scaled_t$`S-181218-03143`)

soma_selection <- soma_scaled_t %>% 
  select(somaid,all_of(samples))

names(soma_selection)[1] <- "marker"



###################################################
# generate a dataframe of pathways and somamers
###################################################

pathway_data <- data.frame(matrix(ncol = 2,nrow = 0))
names(pathway_data) <- c("marker","pathway")
  
for(i in 1:length(pathways)){
    # i <- 1
    working_data <- newcluster_results[,c("seqid",pathways[i])]
    working_data <- working_data[which(working_data[[pathways[i]]]==1),]
    names(working_data) <- c("marker","pathway")
    working_data$pathway <- pathways[i]
  pathway_data <- rbind(pathway_data,working_data)  
  }

# load correlation of somamers with GMSdis outcome

  corrs <- read_csv("./input/ms_vs_hd_forIPA_selection.csv")


pathway_data <- merge(pathway_data,corrs,by="marker",all.x = TRUE)



# 
# get a dataframe of pathways for progression score

results <- tibble::enframe(x = samples,name = NULL,value = "sampleid")

nr_soma <- vector()
cor_dir_sum <- vector()

# generate a file for supplementary table - summary of pathways  

summary_pathways <- data.frame(matrix(nrow = 0,ncol = 6))

names(summary_pathways) <- c("pathway","type","method","seqid","gene","coefficient")

for(j in 1:length(pathways)){
# j <- 1
  
selection <- pathway_data %>% 
    filter(pathway==pathways[j])
  
  pathways_data <- selection %>% 
    dplyr::select(marker,gene,ms_hd_median_diff)

  # merge with seqid data
  
  data_soma <- merge(pathways_data,soma_selection, by="marker",all.x = TRUE)
  
 
  final_data <- data_soma
  
  final_data[,samples] <- final_data$ms_hd_median_diff * final_data[,samples]
  
  nr_somamers <- length(final_data$marker)
  
  # calculate the score
  
  score <- apply(final_data[,samples],MARGIN = 2,FUN = sum, na.rm=TRUE)
  # score <- apply(final_data[,samples],MARGIN = 2,FUN = mean, na.rm=TRUE)
  
  # mutliply score by sign of the sum of correlation coefficient for the pathway
  
  cor_dir <- ifelse(sum(final_data$ms_hd_median_diff)>0,yes = 1,no = -1)
  
  cor_dir_sum <- c(cor_dir_sum,sum(final_data$ms_hd_median_diff))
  
  score <- score * cor_dir
  
  # score <- score / nr_somamers
  results[[paste0(pathways[j],"_score")]] <- score  
  
  # selected somamers
  
  nr_soma <- c(nr_soma,nr_somamers)
  
  #for summary
  for_summary <- final_data %>% 
    mutate(pathway=pathways[j]) %>% 
    mutate(type="ms vs hd") %>% 
    mutate(method="clustering") %>% 
    dplyr::select(pathway,type,method,marker,gene,ms_hd_median_diff) %>% 
    dplyr::rename(somaid=marker) %>% 
    dplyr::rename(coefficient=ms_hd_median_diff)
  
  summary_pathways <- rbind(summary_pathways,for_summary)

      
}

nr_soma_for_scores <- tibble::enframe(pathways,name = NULL,value = "pathway")

nr_soma_for_scores$nr_soma <- nr_soma

write_csv(nr_soma_for_scores,"./output/MS_vs_HD/new_nr_soma_in_scores.csv")




write_csv(results,"./output/MS_vs_HD/new_ms_vs_hd_scores.csv")


# save summary pathways 

write_csv(summary_pathways,"./output/MS_vs_HD/new_ms_vs_hd_summary_pathways.csv")
```
```

